---
title: "Modeling Gender Interests"
output: html_notebook
---


```{r}
library(car)
library(GGally)
library(lime)
library(caret)
library(readr)
library(rpart)
library(d3heatmap)
library(RColorBrewer)
library(dplyr)
library(ggplot2)
library(GGally)
library(corrplot)
library(randomForest)
library(data.table)
library(highcharter)
library(tidyr)
library(factoextra)
setwd("~/Documents/Rscripts /Stat 295/Final")
youngPeople=read.csv("responses.csv", stringsAsFactors = F)
variables = names(youngPeople)
names(youngPeople)[2] = "song speed"
names(youngPeople)[10] = "Metal/HardRock"
#youngPeople$Gender = as.factor(youngPeople$Gender)
youngPeople$Alcohol[youngPeople$Alcohol == "never"] = 1
youngPeople$Alcohol[youngPeople$Alcohol == "social drinker"] = 2
youngPeople$Alcohol[youngPeople$Alcohol == "drink a lot"] = 3
youngPeople$Alcohol = as.numeric(youngPeople$Alcohol)
youngPeople$Smoking[youngPeople$Smoking == "never smoked"] = 1
youngPeople$Smoking[youngPeople$Smoking == "tried smoking"] = 2
youngPeople$Smoking[youngPeople$Smoking == "former smoker"] = 3
youngPeople$Smoking[youngPeople$Smoking == "current smoker"] = 4
youngPeople$Smoking = as.numeric(youngPeople$Smoking)
youngPeople$Internet.usage[youngPeople$Internet.usage == "no time at all"] = 1
youngPeople$Internet.usage[youngPeople$Internet.usage == "less than an hour a day"] = 2
youngPeople$Internet.usage[youngPeople$Internet.usage == "few hours a day"] = 3
youngPeople$Internet.usage[youngPeople$Internet.usage == "most of the day"] = 4
youngPeople$Internet.usage = as.integer(youngPeople$Internet.usage)
youngPeople$Punctuality[youngPeople$Punctuality == "i am often early"] = 3
youngPeople$Punctuality[youngPeople$Punctuality == "i am always on time"] = 2
youngPeople$Punctuality[youngPeople$Punctuality == "i am often running late"] = 1
youngPeople$Punctuality[youngPeople$Punctuality == ""] = NA
youngPeople$Punctuality = as.numeric(youngPeople$Punctuality)
youngPeople$Lying[youngPeople$Lying == "never"] = 1
youngPeople$Lying[youngPeople$Lying == "only to avoid hurting someone"] = 2
youngPeople$Lying[youngPeople$Lying == "sometimes"] = 3
youngPeople$Lying[youngPeople$Lying == "everytime it suits me"] = 4
youngPeople$Lying[youngPeople$Lying == ""] = NA
youngPeople$Lying = as.numeric(youngPeople$Lying)
library("FactoMineR")
```
# Big take aways
1) There is a logistic Regression model that summarizes the relationship between the variables quite well. It holds up to cross validation. (lets write it out in LaTeX)
2) The BIC identifies strong models using almost all variables
3) We should look a bit deeper into this relationship and consider some results from more complex models.
4) lets look further with regularization and tree based methods.

# Next steps
1) We should probably just focus on turning this into the paper, I think it will be quite easy to go from paper to presentation.

2) I do not yet have a baisc tree model in here, It would also be good to fit the GLM model with the glmnet as shown in the lab.  This gives us a bit more information about the coefficents and the graph in the lab does a great job of visualizing the uncertainty around them.

3)Caret's variable importance is genius. We should write up the absolute value t statistic in latex also.

4) Visuals in general are kinda needed at this point for the presentation. But lets focus on the paper first.
https://www.kdnuggets.com/images/cartoon-recommendation-python-machine-learning.jpg
```{r}
split = round(nrow(youngPeople) * .80)
train = youngPeople[1:split, ]
test = youngPeople[(split + 1):nrow(youngPeople), ]
interests = train %>% select(Gender,History:Pets) %>% filter(Gender != "") %>% drop_na() %>% mutate(Gender = as.factor(Gender))
interestsTest = train %>% select(Gender,History:Pets) %>% filter(Gender != "") %>% drop_na() %>% mutate(Gender = as.factor(Gender))
```
### lets start here by graphing that PCA again just as a point of reference
```{r, fig.height=10, fig.width=10}
h=PCA(interests[,2:ncol(interests)], scale = T, graph = F) 
fviz_pca_biplot(h,geom = "point",
                repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                habillage = interests$Gender,
                col.ind = "#696567" # Individuals color
                )
  
```
## Lets try some best subset regression to see what this tells us. I like the regsubsets default better than ISLR's suggestion of using all variables  
```{r}
library(leaps)
regfit <- regsubsets(Gender ~ ., interests, method = "forward")
```
```{r, fig.height=10, fig.width= 10}
par(mfrow=c(2,2))
plot(regfit, scale = 'r2')
plot(regfit, scale = 'adjr2')
plot(regfit, scale = 'Cp')
plot(regfit, scale = 'bic')
```
## Here's what it we get following the books recomendation of using all variables. There's something interesting happening with the BIC
```{r}
library(leaps)
regfit2 <- regsubsets(Gender ~ ., interests, method = "forward", nvmax = 33)
```
```{r, fig.height=11, fig.width=10}
par(mfrow=c(2,2))
plot(regfit2, scale = 'r2')
plot(regfit2, scale = 'adjr2')
plot(regfit2, scale = 'Cp')
plot(regfit2, scale = 'bic')
```



## create folds to prevent leakage
```{r}
myFolds <- createFolds(interests$Gender, k = 10)
set.seed(2017)
```
## Created train control, makes sure that our folds are constant across all models.  Prevents leakage!
```{r}
myControl <- trainControl(
  method = "cv",
  number = 10,
  summaryFunction = twoClassSummary,
  classProbs = TRUE, # IMPORTANT!
  index = myFolds
  
)
pred = function(model){
preds = predict(model, interestsTest)
confusionMatrix(interestsTest$Gender, preds)}

```
## The PCA was able to help us find an interesting dimension in the data now lets dig a little deeper.
## We'll chose this subset of variables based on our PCA and best subset results.

# This really should be our final model.  It would be nice if we could write this up in latex, I don't know how to do that.
https://stats.stackexchange.com/questions/316641/what-is-the-usefulness-of-detection-rate-in-a-confusion-matrix
```{r}
logreg<- train(Gender ~ PC + Reading + Cars + Shopping, interests, method = 'glm',
              trControl = myControl)
logreg
pred(logreg)
summary(logreg)
ggcoef(logreg$finalModel,  errorbar_color = "blue",
  errorbar_height = .25, exclude_intercept = T) + ggtitle("Initial Logistic Regression Model")
vif(logreg$finalModel)
```
## quick note on metrics https://stats.stackexchange.com/questions/124001/what-is-the-intuition-behind-the-kappa-statistical-value-in-classification


```{r}
Tree <- train(
  Gender ~ .,
  tuneLength = 10,
  importance = "impurity",
  data = interests, method = "ranger",
  trControl = myControl)
pred(Tree)
Tree
```
# as far as tree based methods go it seems that extra trees is doing really well. 
https://www.quora.com/What-is-the-extra-trees-algorithm-in-machine-learning
```{r}
treeinfo = as.data.frame(Tree$results)
ggplot(treeinfo, aes(mtry, ROC, color = splitrule))+
         geom_point()+
        ggtitle("Tree Rules")
```
```{r, warning = FALSE}
IMP= function(model){
importance = as.data.frame(varImp(model)$importance) %>% add_rownames("Question") 
#%>% mutate(Question = as.factor(Question))
importance = arrange(importance, desc(Overall)) %>% mutate(Question = as.factor(Question))
order = importance$Question
importance$Question = factor(importance$Question, levels = order)
return(importance)
}

```
```{r}
TreeImportant= IMP(Tree)
```

```{r}
implot= function(importance){
ggplot(importance, aes(Question, Overall))+
  geom_bar(stat = "identity")+
  coord_flip()}
a=TreeImportant %>% implot + ggtitle("Random Forest")
a
```
```{r}
logreg2 = train(Gender ~ Cars + PC + Shopping + Reading + Dancing, interests, method = 'glm',
              trControl = myControl)
pred(logreg2)
```
Notes of Caret Variable Importance: http://ftp.uni-bayreuth.de/math/statlib/R/CRAN/doc/vignettes/caret/caretVarImp.pdf

Here’s more info on the extra trees algorithm https://www.quora.com/What-is-the-extra-trees-algorithm-in-machine-learning
## What This is doing is crazy, this is really one of the best estimate of out of sample error possible with the data.
```{r}
library(lars)
set.seed(825) # for reproducing these results

ridge <- train(Gender ~., data = interests,
               method='glmnet', 
               preProcess=c('scale', 'center'),
               trControl = myControl,
               tuneGrid = expand.grid(alpha = seq(.05, 1, length = 15),
                                               lambda = c((1:5)/10)))
ridge$resampledCM
ridge$resample
ridge

ridgeImportance = IMP(ridge)
b=ridgeImportance %>% implot()+ ggtitle("Lasso Regression")
pred(ridge)
ridge
varImp(ridge, scale = F)
f= youngPeople %>% select(PC, Cars)
```

```{r}
pred(ridge)
```
## This is a really cool idea, these are scaled absolute value t-statistics. Were able to get some sort of way to compare what the two models are saying.
```{r}
fullmodel = train(Gender ~., data= interests, method = 'glm',
              trControl = myControl)
c=IMP(fullmodel) %>% implot() + ggtitle("Logistic Regression")
pred(fullmodel)
varImp(logreg, scale =FALSE )
```
```{r, fig.height= 12, fig.width=14}
library(gridExtra)
grid.arrange(a,b,c) 
```

